
###Homework
```{r}
library(lubridate)
library(mosaicData)
library(tidyverse)
data("Birthdays")
```

```{r}
Birthdays1<-
  Birthdays %>%
  mutate(month=month(date,label=TRUE),
        weekday=wday(date,label=TRUE))
```
###Exercise 5.1
```{r}
septTopFive<-
  Birthdays%>%
  filter(date>="1979-09-09", date<="1979-11-09")%>%
  group_by(state)%>%
  summarise(total_births=sum(births))%>%
  arrange(desc(total_births))
head(septTopFive)
```


###Exercise 5.2

```{r}
DailyBirths<-
  Birthdays%>%
  group_by(date)%>%
  summarise(n=sum(births))
head(DailyBirths)
ggplot(DailyBirths, aes(y=n,x=date))+geom_point()
```

Check by week

wDailyBirths<-
  Birthdays%>%
  group_by(wday)%>%
  summarise(n=sum(births))
head(wDailyBirths)
ggplot(DailyBirths, aes(y=n,x=wday))+geom_point()


Check by month
```{r}
mBirths<-
  Birthdays%>%
  group_by(month)%>%
  summarise(n=sum(births))
head(mBirths)
ggplot(mBirths, aes(y=n,x=month))+geom_point()
```

```{r}
JulianBirths <- Birthdays%>%
  mutate(JulDay=yday(date))%>%
  group_by(JulDay) %>%
  summarise(total=sum(births)) 
ggplot(data=JulianBirths,aes(x=JulDay,y=total))+geom_point()
```

###Exercise 5.3 
```{r}
ggplot(Birthdays, aes(y=births, x=wday))+geom_boxplot()
```

###Exercise 5.4 
```{r}
MyTwoYears<-
  Birthdays%>%
  filter(year==c(1980,1981))%>%
  group_by(date)%>%
  summarise(total=sum(births))
head(MyTwoYears)
```

```{r}
ggplot(MyTwoYears, aes(x=date,y=total, color=wday(date, label=TRUE)))+geom_point()

```

###Exercise 5.5 

```{r}
Holidays <- read_csv("https://tiny.cc/dcf/US-Holidays.csv") %>%
  mutate(date = as.POSIXct(lubridate::dmy(date)))
head(Holidays)
```
a) 
```{r}
MyTwoYears2<-
  MyTwoYears%>%
  mutate(is_holiday = date %in% Holidays$date)
head(MyTwoYears2)
```
b)
```{r}
X<-ggplot(MyTwoYears2, aes(x=date,y=total, color=wday(date, label=TRUE),shape=is_holiday))+geom_point()
X
```
c) 
```{r}
Holidays2<-
  Holidays%>%
  filter(year(date)==c(1980,1981))
head(Holidays2)
```

```{r}
ggplot(MyTwoYears2, aes(x=date,y=total, color=wday(date, label=TRUE)))+geom_point()+geom_text(data=Holidays2, aes(x=date, y=4000, label=holiday),angle=90)

```
###Exercise 5.6

Add a new variable for the top coldest states:
```{r}
BirthdaysCold<-
  Birthdays%>%
  filter(year==c(1980,1981))%>%
  mutate(topCold=state=='MN'|state=='AK'|state=='ND'|state=='ME'|state=='WY'|state=='MT'|state=='VT'|state=='WI'|state=='NH')
head(BirthdaysCold)
```

```{r}
ggplot(BirthdaysCold, aes(y=births, x=date, color=topCold))+geom_point()

```

We can see that the coldest states in the US have lower birth rates. 

###Exercise 5.7
```{r}
library(fivethirtyeight)
USB<-data(US_births_1994_2003)
head(US_births_1994_2003)
```

```{r}
allyears <- rbind(US_births_1994_2003, US_births_2000_2014)
thirteen <- allyears %>%
        filter(date_of_month == 13|date_of_month == 6|date_of_month == 20)
head(thirteen)
```


```{r}
day620<-
  thirteen%>%
  filter(date_of_month == 6|date_of_month == 20)%>%
  group_by(day_of_week)%>%
  summarise(mean620=mean(births))
head(day620)




```

Create mean difference
```{r}
difference<-
  thirteen%>%
  filter(date_of_month == 13)%>%
  group_by(day_of_week)%>%
  summarize(mean13=mean(births))
head(difference)
```

Join on day_of_week

```{r}
averagesTable<-
  day620%>%
  full_join(difference)
head(averagesTable)
```

Create a table with the differences in percentages
```{r}
finalTable<-
  averagesTable%>%
  group_by(day_of_week)%>%
  summarise(difference=((mean13-mean620)/mean620))
head(finalTable)
```

Create the plot
```{r}
library(ggplot2)
```

```{r}
ggplot(finalTable, aes(x=day_of_week,y=difference))+geom_col(fill="pink")

``` 


